---
title: "Task 1 - California Oil Spills"
author: "Max Settineri"
date: "2023-03-18"
output: 
  html_document:
    code_folding: hide
---

## Overview


**Oil Spill Data Citation:** Lampinen, Mark (2009). Oil Spill Incident Tracking [ds394]. California Department of Fish and Game, Office of Spill Prevention and Response.
https://map.dfg.ca.gov/metadata/ds0394.html?5.108.39

**California County Data Citation:** 

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
library(tmaptools)
library(terra)
```

## Wrangling oil spill and county data

```{r}
## reading in oil spill data
oil <- read_csv(here("data/Oil_Spill_Incident_Tracking_[ds394].csv")) %>% 
  clean_names() 
  
## reading in county shapefile data and keeping only county name and geometry
counties <- read_sf(here("data/CA_Counties_TIGER2016.shp")) %>% 
  clean_names() %>% 
  select(name)

## converting oil spill data to sf
oil_sf <- st_as_sf(oil, coords = c("x", "y"),
                   crs = st_crs(counties)) # setting oil crs to match counties data

## checking to make sure crs matches
# st_crs(counties)
# st_crs(oil_sf)

```

## Interactive map of California oil spill locations

```{r}
## setting tmap mode and map style:
tmap_mode("view")
tmap_style("natural")

## mapping oil spill occurrences by CA county
tm_shape(counties) +
  tm_polygons(alpha = 0) +
tm_shape(oil_sf) +
  tm_dots(col = "salmon") 
```

**Figure 1:** Here we can observe the locations of oil spill incidents in the state of California. Red dots indicate locations of California oil spill occurances.  

## Choropleth map of inland oil spill locations in California

```{r}
## including only inland oil spills
oil_inland <- oil_sf %>% 
  filter(inlandmari == "Inland")

## joining oil spill data to counties data
oil_inland_county <- counties %>% 
  st_join(oil_inland) %>% 
  ## getting counts of spills by county
  group_by(name) %>% 
  summarize(spills = n())

## creating choropleth map of spill counts by county
ggplot(data = oil_inland_county) +
  geom_sf(aes(fill = spills)) +
  theme_void() +
  scale_fill_gradientn(colors = c("slategray1", "purple4"))
```

**Figure 2:** This figure displays inland oil spill frequency by California county. Darker purple indicates a higher number of oil spills, while lighter colors indicate a lower number of spills. 